On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence

### 2017-09-18 Thesis/eigenstatement.Rmd
Printed version

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=Blue,
            urlcolor=black,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{}
  \pretitle{\vspace{\droptitle}}
  \posttitle{}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage[lined,boxed]{algorithm2e}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newcommand{\KL}{\,\text{KL}}
\newcommand{\der}{\,\text{d}}
\newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%

\begin{document}

\pagenumbering{gobble}

\begin{centering}

\vspace{5cm}

    {\Large Humboldt-Universit\"at zu Berlin \\
    School of Business and Economics \\
    Institute for Statistics and Econometrics\\}

    \vspace{4cm}

{\huge{\bf On Pure Exploration, Thresholding Bandits, \\
and Kullback-Leibler Divergence\\}}

\vspace{2cm}

{\large Master's Thesis submitted\\\vspace{0.5cm}
    to}
    \\\vspace{0.5cm}
    {\large{\bf Prof. Dr. Marius Kloft \\}}
    {\large{\bf Prof. Alexandra Carpentier, PhD \\}}
    \vspace{0.5cm}

    {\large by \\\vspace{0.5cm}
    {\bf Tim Radtke\\}
    (572481)\\
    born on 11.01.1992 in Viersen \\\vspace{1cm}}

    {\large in partial fulfillment of the requirements \\
    for the degree of \\
    {\bf Master of Science (M. Sc.)\\} \vspace{3cm}
    Berlin, September 18, 2017\\}

\end{centering}

\newpage

\vspace*{10\baselineskip}

``In general, we look for a new law by the following process: First we
guess it. Then we -- now don't laugh, that's really true. Then we
compute the consequences of the guess to see what, if this is right, if
this law that we guessed is right, to see what it would imply. And then
we compare the computation results to nature, or we say compare to
experiment or experience, compare it directly with observations to see
if it works. If it disagrees with experiment, it's wrong. In that simple
statement is the key to science. It doesn't make any difference how
beautiful your guess is, it doesn't make any difference how smart you
are, who made the guess, or what his name is. If it disagrees with
experiment, it's wrong. That's all there is to it.''
\vspace*{1.5\baselineskip}

Richard Feynman in \emph{Richard Feynman Messenger Lectures at Cornell
University: The Character of Physical Law. Chapter 7, Seeking New Laws}.

\newpage

\vspace*{10\baselineskip}

\Large{\bf Abstract\\}

\vspace*{1\baselineskip}

\normalsize
In this thesis we present a new sequential sampling strategy for the
fixed-budget thresholding bandit problem described in Locatelli et al.
(2016). We draw from the literature on pure exploration bandit problems
to show the similarity of the thresholding bandit to (sequential)
hypothesis testing. In particular, we derive a lower bound on how fast
the error probability in a one-armed thresholding bandit problem can
decrease at best, and show that this corresponds to the Chernoff-Stein
Lemma. This motivates the use of the Kullback-Leibler divergence in the
index of our new algorithm. In simulations, our strategy is shown to
outperform existing strategies. Using website analytics data from an
online shop, we use time series cross-validation to create data sets for
offline testing of adaptive sampling strategies; time dependency in the
data is thus preserved. Our experiments show that the adaptive sampling
strategies are disrupted by the dependency structure in the samples.

\newpage

\setcounter{tocdepth}{2} \tableofcontents

\newpage

\pagenumbering{arabic}

\section{\texorpdfstring{Introduction
\label{chap:Introduction}}{Introduction }}\label{introduction}

The thesis at hand joins a large literature concerned with the problem
of sequential sampling. One is familiar with the idea of ``batch
sampling'', where the entire sample is gathered first and analyzed only
after the last sample has become available. However, as noted in
Chernoff (1958), many situations arise in which one would like to gather
feedback from samples as they are drawn, so as to influence the choice
of the next sample. The two fundamental characteristics of the latter
sampling process are that it is sequential and adaptive. Since
observations are drawn in sequence, one has a chance to analyze past
observations in order to adapt the decision of what sample one would
like to observe next.

This makes sequential sampling especially relevant in situations where
there is an inherent cost to obtaining samples, or where only a limited
amount of samples can be collected. The classical example is that of
clinical trials in which a researcher tests two drugs on patients to
observe which one performs better. Not only is it difficult to gather
enough patients, but one would also like to adapt the sampling such that
more patients receive the better drug as soon as possible, and are
cured. As noted in Cappé et al. (2013), this example dates back to
Thompson (1933, 1935). Such problems are known as multi-armed bandit
problems in which the sampling is adapted so as to minimize the
cumulative regret: In the case of the drug tests, one would like to
sample the two drugs in a way that quickly finds the best drug and then
tests this drug more and more often, and more often than the worse drug.
This would lead to cumulatively more healthy patients than testing both
drugs uniformly; it would minimize the regret from treating patients
with an ineffective drug.

Instead of drugs, one can consider two or more unknown distributions
(which we call arms in the multi-armed bandit problem in analogy with
the one-armed bandit slot machine). These distributions have different
means, and one can draw samples sequentially from the distributions in
order to estimate their means. To minimize the cumulative regret
measure, one would like to sample from the distribution with the highest
mean as often as possible (\emph{Exploitation}). At the same time,
however, one has to sample all arms at least a little, so as to
initially find the best arm (\emph{Exploration}). Any algorithm thus has
to handle a trade-off between exploration and exploitation to minimize
the regret. How can we organize the sampling in order to achieve this
task?

While Gittins (1979) presents an Bayesian-optimal solution to the
cumulative regret problem, the Upper-Confidence Bound (UCB) algorithms
pioneered by Agrawal (1995) and Auer, Cesa-Bianchi and Fischer (2002)
have lead to a blooming literature on multi-armed bandit problems that
has been picked up to solve modern problems in website and advertising
optimization, where advertisers try to allocate their advertising
optimally so as to reduce regret from showing poorly converting
advertisements. Several algorithms like the KL-UCB algorithm based on
Kullback-Leibler divergence upper confidence intervals (Cappé et al.,
2013) or the Bayes-UCB algorithm (Kaufmann et al., 2012a) have followed
and been shown to be optimal as they reach the lower bound introduced in
Lai and Robbins (1985) and Burnetas and Katehakis (1996).

These algorithms have in common that they compute after every new sample
an index for each of the unknown distributions. This index is based
solely on the past (observed) data, and then used to pick the
distribution from which the next sample should be drawn. Using these
indexes, algorithms can sample sequentially and adapt their strategies
to observations as they are collected.

Similar algorithms based on indexes can be used on related problems.
Instead of minimizing cumulative regret, it can be of interest to find
the best arm, that is, the distribution with the largest mean -- with as
few samples as possible and with confidence as large as possible. This
is particularly useful when there is not a cost from testing arms with
lower mean, but when drawing any sample is costly. Thus one needs to
explore the distributions as efficiently as possible. Given that the
loss function only evaluates the classifiation once at the end of the
sampling process, exploitation is \emph{not} necessary, and the setting
is thus referred to as \emph{Pure Exploration}. The need for strategies
different from uniform sampling arises still: One does not want to
continue to sample distributions that one is already confident in. This
setting is related to the problem faced in Bayesian Optimization:
Training of, for example, deep neural network architectures is costly in
terms of time and computation, and so the parameter space needs to be
explored with as few samples as possible. (Zhang et al., 2017; Brochu et
al., 2010). The problem of best arm identification is also very related
to sequential statistical hypothesis testing as noted by Chernoff
(1958), and picked up by Garivier and Kaufmann (2016) and Russo (2016),
since the experimenter has to make a decision about how the means of
distributions relate to each other.

Given the focus on exploration, the pure exploration problem is usually
described with a loss function different from the one in the cumulative
regret settings. Audibert et al. (2010) introduce a soft measure called
\emph{simple regret} based on which they describe UCB policies. Other
authors investigate binary measures that are equal to 1 when the best
arm has not been identified (Russo, 2016; Garivier and Kaufmann, 2016).
These binary measures correspond to the probability of the identified
arm being the best, and thus are inherently related to hypothesis
testing. The problem of best arm identification can readily be extended
to the so called Top-\(m\) problem of identifying the best \(m\) of
\(K\) arms (Kaufmann and Kalyanakrishnan, 2013; Kaufmann et al. 2016).

In this thesis, we consider the thresholding bandit problem. While the
name is reminiscient of the cumulative regret multi-armed bandit, the
problem is more similar to the top-\(m\) problem. But instead of
identifying the best \(m\) out of \(K\) arms, one aims to find all
distributions out of \(K\) that have a mean at least as large as a fixed
threshold (it is not known upfront how many distributions fulfill this
condition). Given that an implicit threshold arises in the top-\(m\)
problem between the \(m\)-th best and the \((m+1)\)-th best arm
(Kaufmann et al., 2016), the two problems share many characteristics. We
discuss this in more detail in
\autoref{chap:KullbackLeiblerPureExploration}. In the setting described
in Locatelli et al. (2016), algorithms in the thresholding bandit
problem are evaluated against a binary loss function penalizing once
after the sampling process if an arm has not been classified correctly
as being above or below the threshold. This setting thus focuses on
exploration similar to algorithms employed for best arm identification.
Indeed, Locatelli et al. (2016) show how their thresholding algorithm
can be used for best arm identification when the mean of the best arm is
known upfront.

In the following, we first present the thresholding bandit problem in
detail in \autoref{chap:ThresholdingBanditProblem}. To then extend upon
the algorithm proposed by Locatelli et al. (2016) and discussed in
\autoref{sec:TheAPTAlgorithm}, we first explore in
\autoref{chap:KullbackLeiblerPureExploration} how the Kullback-Leibler
divergence has been employed in the best arm identification literature
by Garivier and Kaufmann (2016) and Kaufmann et al. (2016) as this will
lead us to our own extension. We shortly describe in
\autoref{sec:VarianceBasedAlgorithms} two variance-based algorithms for
the thresholding bandit problem introduced in Mukherjee et al. (2017)
and Zhang et al. (2017), before introducing the new Simple Likelihood
Ratio algorithm in \autoref{sec:LRforUnivariateExponentialFamily}. We
end in \autoref{chap:Experiments} with experiments to examine the
performance of different algorithms.

\newpage

\section{\texorpdfstring{The Thresholding Bandit Problem
\label{chap:ThresholdingBanditProblem}}{The Thresholding Bandit Problem }}\label{the-thresholding-bandit-problem}

\subsection{An Example}\label{an-example}

To introduce the thresholding bandit problem, consider a common
situation among internet-based companies. These companies often collect
user feedback to offer reviews and ratings on websites or in apps, or
collect it to improve their services. Netflix lets users rate movies
with ``Thumbs up'' and ``Thumbs down'', apps like Clue ask users a
question per day to learn to predict the user's health and mood. An
online shop might give customers a one-question survey after they make a
purchase. Finally, Foursquare and Yelp ask users to rate coffee shops,
bars, and restaurants.

Consider the problem the latter companies face. Whenever a user reports
that she visited a restaurant, the app would like to ask questions about
her experience. To give a simple example, assume users are not willing
to answer more than one question per restaurant visit, and the app would
like to find out two things: Does the restaurant have good service? Is
the restaurant trendy? See \autoref{app:Foursquare} for how this
information is then displayed in the Foursquare app.

If the feedback that users can give is restricted to binary ``Yes'' or
``No'' answers, then we can model the feedback as Bernoulli random
variables

\[
X_{\text{service}} \sim \text{Ber}(\mu_{\text{service}}) \quad \text{and} \quad X_{\text{trendy}} \sim \text{Ber}(\mu_{\text{trendy}}).
\]

The app will display a ``Good service'' or ``This place is trendy''
description for the restaurant if \(\mu_{\text{service}}\) and
\(\mu_{\text{trendy}}\) are estimated to be above a certain threshold,
for example \(\tau = 0.5\). That is, more than half of the user base
would describe the place as having good service, or as trendy.

To make the app as useful as possible to other users, this information
should be shown as quickly as possible. But even though the app has
thousands of users, not that many people visit a specific restaurant,
and even less people report this back to the app. Thus the company has
only a limited budget of \(T\) potential users from which to collect
samples \(X_{\text{service}}\) and \(X_{\text{trendy}}\). How can the
company ask these \(T\) questions to be as confident as possible in its
classification of the restaurant's features, that is, in the information
it displays?

A straightforward way of sampling users would be to ask the two
questions in a round-robin fashion: Ask for ``Service'', then for
``Trendy'', then for ``Service'', \ldots{} Given 100 users, 50 users
give feedback on the service level, 50 users report on the trendiness
level. This, however, is unlikely to be optimal with respect to the
confidence after 100 samples. Given that the company asks the users
sequentially, there is the option to adapt the sampling strategy over
time: If the first ten users report
\(\sum_{s=1}^{5}X_{\text{service},s}=5\), while
\(\sum_{s=1}^{5}X_{\text{trendy},s}=3\), then the company might already
be quite confident that \(\mu_{\text{service}} \geq \tau = 0.5\), while
this is less clear for \(\mu_{\text{trendy}}\). Instead of continuously
sampling both questions, it might be more useful to collect more
feedback on the trendiness level so as to ensure its position relative
to the threshold. After all, the company would like to use the 100
samples to maximize its confidence in \emph{both} displayed tips.

Finding the optimal sequential sampling strategy to classify
distributions against a threshold with as much confidence as possible
after a limited budget of samples is the problem described above. It is
a setting described by the thresholding bandit problem, which we discuss
for the remainder of this thesis. In the following section, we define
setting of the thresholding bandit problem more generally.

\subsection{\texorpdfstring{General Setting
\label{sec:Setup}}{General Setting }}\label{general-setting}

Consider a setting with unknown probability distributions \(\nu_i\)
indexed by \(i \in \mathbb{A}=\{1, ..., K\}\). To stick with the idea of
multi-armed bandits, we call the distributions arms, and observations
from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is
interested in learning about the structure of the arms; in particular,
it would like to estimate the means \(\mu_1, ..., \mu_K\) of each
distribution \(\nu_i\) by sequentially drawing observations
\(X_{i,1}, X_{i,2}, ...\) from the different arms, where
\(X_{i} \sim \nu_i\). However, at time \(t\) the learner cannot observe
a reward from every distribution but must select a single arm \(I_t\)
from which it would like to observe the reward \(X_{I_t,t}\). The
decision of which arm to pull next is often made by computing an index
\(B_i(t)\) for each arm. If \(T_i(t)\) is the number of samples observed
until time \(t\) for arm \(i\), then \(B_i(t)\) is a function of the
past rewards \(X_{i,1}, ..., X_{i,T_i(t)}\) observed for arm \(i\), and
w.l.o.g. \(I_t = \arg \min_{i = 1,...K} B_i(t)\). The set of indexes can
be called strategy (Kaufmann et al., 2012a) which makes it clear that
the strategy is adapted as more observations are collected. Strategies
of this kind are also called index policies.

Which index proves to be useful is determined by the loss function with
which the decisions of the algorithm are evaluated. Many problems are
concerned with finding the arm with the largest mean \(\mu_i^*\) while
collecting loss either cumulatively during (Auer and Cesa-Bianchi, 2012)
or once at the end of the sampling (Kaufmann et al., 2016). The
thresholding bandit problem considers a slightly different loss
function. Consider a fixed threshold value \(\tau \in \mathbb{R}\) and
an epsilon interval \(\tau \pm \epsilon\) around the threshold for
\(\epsilon \geq 0\). As mentioned in the introduction, the learner is
tasked with classifying arms as having a mean \(\mu_i\) larger or
smaller than \(\tau\). As in Locatelli et al. (2016), whenever we say
that ``an arm is over the threshold'', we say that the mean of the arm
is over the threshold. To describe which arms are over the threshold, we
also introduce the following set notation: Let \(\mathcal{S}_{u}\) be
the set of distributions in \(\mathbb{A}\) with mean larger than some
\(u \in \mathbb{R}\):
\(\mathcal{S}_{u} := \{i \in \mathbb{A}, \mu_i \geq u\}\). Consequently,
the complement is given by the arms with mean below \(u\),
\(\mathcal{S}^C_{u} := \{i \in \mathbb{A}, \mu_i < u\}\).

The thresholding problem for the learner is now to use a budget of
overall \(T\) samples (or rounds) to explore the different arms. After
this exploration phase, the learner returns a set of arms that it
believes to be over the threshold. When evaluating the classification of
the learner, arms with mean inside of the epsilon interval
\(\tau \pm \epsilon\) are not considered. The learner only has to
correctly distinguish the means from the threshold up to the precision
given by \(\epsilon\). Thus, the algorithm is not evaluated against
\(\mathcal{S}_{\tau}\) and \(\mathcal{S}^C_{\tau}\), but needs to
distinguish \(\mathcal{S}_{\tau+\epsilon}\) from
\(\mathcal{S}^C_{\tau-\epsilon}\).

The learner's classification is returned as the set
\(\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) \subseteq \mathbb{A}\)
after \(T\) rounds; that is, the set of arms estimated to be over the
threshold. The classification is evaluated against the binary loss given
by (Locatelli et al., 2016):

\[
\mathcal{L}(T) = \mathbbm{1}{\big((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor (\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)\big)}
\]

In order to perform well, a learner has to identify all arms that are
over the threshold as such, and to not claim arms as being over the
threshold if they actually are below the threshold. Due to the precision
\(\epsilon\), the loss is not influenced by arms which are very close to
the threshold \(\tau\) as measured by \(\epsilon\), that is, arms inside
of the \(\tau \pm \epsilon\) interval. If either
\(\hat{\mathcal{S}}_{\tau}\) contains an arm below
\((\tau - \epsilon)\), or \(\hat{\mathcal{S}}^C_{\tau}\) contains an arm
above \((\tau + \epsilon)\), a loss of \(1\) is accrued. Else, all arms
outside the epsilon interval are classified correctly and the loss is
equal to \(0\). In the case of \(\epsilon = 0\), all arms need to be
classified correctly to not collect any loss.

Algorithms for this thresholding problem can be compared based on their
expected loss. Let \(\mathbb{E}\) be the expectation with regard to the
samples \(X_{i,1}, ..., X_{i,T}\) collected by the algorithm. Then

\[
\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}\big((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor(\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)\big),
\]

is the expected loss, or the probability of making at least one wrong
classification. Both the probability of not identifying an arm that is
actually over the threshold, as well as the probability of accepting an
arm as over the threshold when it is not, are evaluated. The oracle
strategy would evaluate all arms correctly with an error probability of
\(0\) (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit
problem setting which we will use throughout the thesis. The setting can
be extended by different assumptions to make certain algorithms or
bounds feasible. For that reason, Locatelli et al. (2016) assume that
the distributions \(\nu_i, i \in \mathbb{A}\) are \(R\)-sub-Gaussian.
Under this assumption, the authors are able to prove a lower bound on
the largest probability of error of the best algorithm in the setting
just introduced, as well as an upper bound on the performance of their
APT algorithm which we will show in the following chapter. As in their
paper, we define \(R\)-sub-Gaussian as follows:

\begin{definition}[$R$-sub-Gaussian distribution] 
\label{definition:rsubgaussian}
Let $R > 0$. A distribution $\nu$ is $R$-sub-Gaussian if for all $t \in \mathbb{R}$ we have
\begin{equation*}
\mathbb{E}_{X \sim \nu}[\exp(tX - t\mathbb{E}[X])] \leq \exp(R^2t^2/2).
\end{equation*}
\end{definition}

As noted by Locatelli et al. (2016), this ensures that the mean
\(\mu_i\) of the distributions \(\nu_i\) considered in our problem are
finite. \(R\)-sub-Gaussian distributions offer concentration
inequalities for the concentration of the empirical mean around the true
mean, which makes it possible to deduce the bounds presented by the
authors and repeated below. Among other distributions, the family of
\(R\)-sub-Gaussian distributions encompasses Gaussian distributions of
variance \(R^2\) for \(R \in \mathbb{R}\), as well as all bounded
distributions, with the important case of Bernoulli distributions.

Having described the thresholding bandit setting as well as the
assumption of \(R\)-sub-Gaussianity, we now proceed to describe the
\emph{Anytime Parameter-free Thresholding} (APT) algorithm proposed in
Locatelli et al. (2016) for the thresholding bandit problem.

\subsection{\texorpdfstring{The Anytime Parameter-free Thresholding
Algorithm
\label{sec:TheAPTAlgorithm}}{The Anytime Parameter-free Thresholding Algorithm }}\label{the-anytime-parameter-free-thresholding-algorithm}

In the following we describe the \emph{Anytime Parameter-free
Thresholding} (APT) algorithm proposed in Locatelli et al. (2016), which
is a simple yet effective algorithm and displays some very favorable
characteristics when compared to alternative algorithms.

We fix the threshold \(\tau\) and epsilon interval around the threshold
given by \(\epsilon\) upfront, and supply them as arguments to the
algorithm: \((\tau, \epsilon)\). Based on the general thresholding
problem, we demand that the algorithm draws samples \(X_{i,t}\) for
\(T\) rounds from the distributions \(\nu_i, i\in \mathbb{A}\). For a
given arm \(i\) and a given time \(t\), we let \(T_i(t)\) represent the
number of samples drawn from arm \(i\) until time \(t\) (including).
Consequently, the empirical mean of arm \(i\) at time \(t\) is given by
\(\hat{\mu}_i(t) = \frac{1}{T_i(t)} \sum_{s=1}^{T_i(t)} X_{i,s}\) where
\(X_{i,s}\) is the feedback sampled from distribution \(\nu_i\) when the
arm is pulled for the \(s\)-th time. In the case of the APT algorithm,
the decision \(\hat{\mathcal{S}}_{\tau}\) is based on the empirical
means \(\hat{\mu}_i(T)\) estimated from the samples drawn up to time
\(T\):
\(\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) = \{i \in \mathbb{A}, \hat{\mu}_i(T) \geq \tau\}\).

As we see in algorithm \ref{alg:APT}, the APT algorithm initiates by
pulling each arm once. Then, for each \(t > K\), it calculates for every
arm the values \(T_i(t)\) and \(\hat{\mu}_i(t)\). Based on the empirical
mean, the algorithm computes the estimated gap between the mean and the
threshold as
\(\hat{\Delta}_i(t) = \hat{\Delta}_i^{\tau, \epsilon}(t) = |\mu_i(t) - \tau| + \epsilon\).
The decision of which arm is sampled next is then based on the value
\(B_i(t+1) = \sqrt{T_i(t)} \hat{\Delta}_i(t)\): the algorithm pulls arm
\(I_{t+1} = \arg \min_{i\leq K} B_i(t+1)\) minimizing this decision
quantity. After overall \(T\) pulls, the algorithm returns the sets
described in the general setup, \(\hat{\mathcal{S}}_{\tau}\) and
\(\hat{\mathcal{S}}^C_{\tau}\).

\IncMargin{1em}

\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{i \in \{1,...,K\}, \hat{\mu}_i(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{i \in A} B_i^{APT}(t) = \arg \min_{i \in A} \sqrt{T_i(t-1)} \cdot(|\hat{\mu}_i(t-1) - \tau|+\epsilon)$}\;
  \emph{Observe reward $X_t \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{alg:APT}
\end{algorithm}

\DecMargin{1em}

Just like the general index policies mentioned in \autoref{sec:Setup},
we see that the APT algorithm adapts its strategy using an index
\(B_i(t+1)\) for each arm that is based entirely on the observations
observed until time \(t\). Given the arm with the smallest index is
pulled at round \(t+1\), we observe that arms estimated to be further
away from the threshold \(\tau\) as measured by the absolute distance
\(\hat{\Delta}_i(t)\) to their empirical mean \(\hat{\mu}_i\) are pulled
less often. Intuitively, it is easier to classify arms as below or above
\(\tau\) when their mean is far away from the threshold. In contrast to
the uniform sampling strategy, APT will gather more evidence for arms
which appear to be inherently more difficult to classify. Likewise, APT
tends to less often pull arms which it already pulled considerably often
as measured by \(T_i(t)\). If arm \(i\) has been pulled more often than
arm \(j\), then we should in general be more confident in its estimated
mean and the corresponding distance to the threshold. If we are
confident in \(i\), then it is better to explore other arms by drawing
an observation from arm \(j\). By adapting its sampling strategy in this
way during the sequential sampling process, the APT algorithm minimizes
the probability of error \(\mathbb{E}[\mathcal{L}(T)]\) after \(T\)
samples have been observed.

The previous paragraph describes how arms close to the threshold are
intuitively more difficult to classify. In the index employed by the APT
algorithm, the distance between an arm and the threshold is measured by
the absolute distance \(\hat{\Delta}_i(t)\) between the (estimated) mean
of the arm and the threshold \(\tau\) (and the epsilon interval). When
the APT algorithm realizes that an arm has a mean close to the
threshold, then this arm requires more attention than an arm far away
from the threshold; the confidence in the empirical mean has to be
larger. Or the empirical mean has to be better concentrated around the
true mean. Otherwise it is easy to make a mistake on this arm. To gain
this confidence, the arm requires more samples, and thus a larger share
of the overall \(T\) samples available. With this intuition, it makes
sense to measure the complexity of a thresholding problem: If we knew
upfront how complex a given set of arms is, we would know whether to
expect a small or large probability of error given our budget \(T\).
Also, for example in the best arm identification problem, there exist
algorithms which take the complexity of a problem as input. Based on the
complexity they then adapt their sampling strategy. See for example the
UCB-E algorithm in Audibert et al. (2010) for the best arm
identification problem, or the KL-LUCB-E algorithm in Kaufmann and
Kalyanakrishnan (2013) as extension to the top-\(m\) problem.

While the APT algorithm does not require the complexity to be known
upfront (which would be problematic as it is \emph{not} known upfront),
it is necessary to define the complexity to deduce the upper bound on
its probability of error. Similar to the distance measure used in the
algorithm's index, we define the gap \(\Delta_i^{\tau,\epsilon}\)
between arm \(i\) and the threshold \(\tau \pm \epsilon\) as

\[
\Delta_i := \Delta_i^{\tau,\epsilon} = |\mu_i - \tau| + \epsilon
\]

Given this gap, Locatelli et al. (2016) define the complexity
\(H_{\tau, \epsilon}\) for the thresholding bandit problem naturally as

\[
H := H_{\tau, \epsilon} = \sum_{i=1}^K \frac{1}{(\Delta_i^{\tau,\epsilon})^2}.
\]

Note that the complexity of the problem increases in the number of arms
\(K\) and decreases in the gap \(\Delta_i^{\tau,\epsilon}\).
Consequently, when we require less precision, that is, we increase
\(\epsilon\), the hardness decreases. Indeed, by using \(\epsilon > 0\),
we upper bound the complexity of a problem. Consider a problem with
\(K\) arms. Then even in the worst case of \(\mu_i = \tau\) for all
\(i \in \mathbb{A}\), \(H \geq \frac{K}{\epsilon^2}\). Furthermore, we
see that each arm contributes its individual complexity,
\(h_i^{\tau,\epsilon} = \frac{1}{(\Delta_i^{\tau,\epsilon})^2}\)
linearly into the problem. Thus, if the complexity \(H\) is a good
measure of the actual problem complexity, and we knew upfront which arm
has which mean, we could allocate the ``optimal'' share of the budget,
\(T \cdot h_i /H\) to arm \(i\). However, since we know neither
\(\mu_i\) nor \(H\) upfront, the best we can do is to adapt our
allocation of samples more and more towards this optimum as we observe
samples.

At this point, one can question whether the measure of complexity \(H\)
proposed in Locatelli et al. (2016) is a good measure of the ``true''
problem complexity. In the following \autoref{sec:UpperBoundLocatelli},
we will show the upper bound on the probability of error for the APT
algorithm and see how the complexity \(H\) arises naturally for the APT
algorithm. In \autoref{sec:LowerBoundLocatelli} we then show the lower
bound presented in Locatelli et al. (2016), and that the APT algorithm
is optimal in that it matches the lower bound. We discuss the results
and the APT algorithm further in \autoref{sec:DiscussionLocatelli}.

\subsection{\texorpdfstring{Upper Bound for the Anytime Parameter-free
Thresholding Algorithm
\label{sec:UpperBoundLocatelli}}{Upper Bound for the Anytime Parameter-free Thresholding Algorithm }}\label{upper-bound-for-the-anytime-parameter-free-thresholding-algorithm}

In this chapter we present the upper bound on the probability of error
that the APT algorithm incurs, as proven in Locatelli et al. (2016). The
upper bound is on the probability of error that the algorithm has after
\(T\) rounds of sampling for a problem with arms that are assumed to be
\(R\)-sub-Gaussian. The bound depends also on the complexity of the
problem, and we show in the proof that given the index of the APT
algorithm as defined in \autoref{sec:TheAPTAlgorithm}, the complexity
\(H\) is an appropriate measure.

We now state the theorem as in Locatelli et al. (2016) and present a
proof.

\begin{theorem}[Theorem 2, Locatelli et al., 2016] \label{theorem:LocatelliTheorem4}
Let $K \geq 0$, $T \geq 2K$, and consider a problem $\mathcal{B}$. Assume that all arms $\nu_i$ of the problem are R-sub-Gaussian with means $\mu_i$. Let $\tau \in \mathbb{R}, \epsilon \geq 0$. Algorithm APT's expected loss is upper bounded on this problem as 
\begin{equation*} \mathbb{E}(\mathcal{L}(T)) \leq \exp
\Big(-\frac{1}{64R^2}\frac{T}{H} + 2 \log((\log(T) + 1)K)\Big) \end{equation*}

where we remind that $H = \sum_i (|\mu_i - \tau | + \epsilon)^{-2}$ and where
$\mathbb{E}$ is the expectation according to the samples of the problem.
\end{theorem}

\subsubsection{\texorpdfstring{Proof of Theorem
\ref{theorem:LocatelliTheorem4}
\label{sec:UpperBoundProof}}{Proof of Theorem  }}\label{proof-of-theorem}

In order to upper bound the expected loss of the APT algorithm, we want
to define an event on which the algorithm correctly seperates arms from
the threshold, that is, a separation condition holds on this event. We
then want to show that this \emph{favorable} event does indeed hold with
large probability. Consequently, APT's probability of error will
correspond to the probability that the favorable event does not hold, as
this leads to a failure of the seperation condition, and consequently to
an incorrect classification of at least one arm by the algorithm. We
start by defining the separation condition.

\subsubsection{Separation Condition}\label{separation-condition}

As noted in Zhong et al. (2017), a simple requirement to correctly
classify arms at the final round is the following. For every arm
\(i \in \{1, ..., K\}\) after the final round \(T\), require: \[
| \hat{\mu}_{i}(T) - \mu_i | \leq \frac{|\mu_i - \tau| + \epsilon}{2} = \frac{\Delta_i}{2}.
\]

>>>>

---
output: pdf_document
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage{setspace}
---

\pagenumbering{gobble}
\normalsize Copyright © 2017 Tim Radtke \newline
tim@timradtke.net \newline

\vspace*{0.5\baselineskip}

\normalsize Humboldt-Universität zu Berlin \newline
Unter den Linden 6 \newline
10099 Berlin \newline
Germany \newline
